Oregon: Zip-Level Wildfire Smoke and Health Outcomes (open in script)

### 2017-05-05 r_scripts/analysis/or_result_zip_smoke.Rmd
Update or_result_zip_smoke.Rmd
updated to the latest version with plots and tables are readable.

---
title: "Oregon: Zip-Level Wildfire Smoke and Health Outcomes (open in wildfire, not script)"
author: "Jingyang Liu"
date: "April 7, 2017"
output: html_document
---

@@

@@

for (i in var_list){
  read_path <- paste('../data_new/update/ER/or', i, 'may_to_sep_time_strat_casecross.csv', sep = "_")
  or_disease <- read_csv(read_path)

@@

  # Create a permanent case-cross over dataset
  file_name <- paste('../data_new/update/ER/or', i, 'may_to_sep_time_strat_casecross_er.csv', sep = "_")

@@

```{r echo=FALSE, message=FALSE, warning=FALSE}
## Analysis
path <- paste0("../data_new/update/ER")

@@

# read zip pm for counts by zip
smoke_path <- paste0("../data/Oregon/")
zip_pm <- read_csv(paste0(smoke_path, "zip_pm_to_merge_with_chars.csv"))
```

@@

# set list and do cross tabs to find values
# Producing conditional logit model estimates loop 
for(i in 1:length(df_list)){

  i = 2
# dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- colnames(df_to_loop[76])
  # outcome name
  outcome_name <- outcome_list[i]

  # dataframe for analysis creation
  # bind columns back together 
  df_analysis <- df_to_loop %>%
    # only look at outcomes
    filter(outcome == 1) %>%
    # add another row that makes sure there is a person <15 in the dataframe
    # tricking xtabs to produce a 0 cell for the outcome for age <15
    add_row(outcome = 0, age_ind = 0)

@@

### Time-Stratified Case-Crossover (referents same day of week within fire season July - October)

For these analyses and for the final results in the paper, I used referents within the entire wildfire season (July to October). This time-stratified referent strategy allows even more referent observations to average to 'smooth' out some of those extreme PM~2.5~ concentrations in some referent periods. For my sensitivity analysis of varying referent periods, please see other scripts.

Adjusting for CDC Met Variables: Rish has provided zip-level meteorologic data on wind speed, precipitation, relative humidity, and temperature. Note a slight difference that will likely need to be addressed in discussion is that these measures are weighted based on census tract, not on the WRF grid like our measures of smoke.

@@

print(print_plot)
# save figure
ggsave("./fig_only_er.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")
```

@@

>>>>

---
title: "Oregon: Zip-Level Wildfire Smoke and Health Outcomes (open in script)"
author: "Jingyang Liu"
date: "April 7, 2017"
output: html_document
---

@@

### Time-Stratified Case-Crossover (referents same day of week within fire season May - September)

@@

for (i in var_list){
  read_path <- paste('../data_new/casecrossover/time_strat/or', i, 'may_to_sep_time_strat_casecross.csv', sep = "_")
  or_disease <- read_csv(read_path)

@@

  # Create a permanent case-cross over dataset
  file_name <- paste('../data_new/casecrossover/ER/or', i, 'may_to_sep_time_strat_casecross_er.csv', sep = "_")

@@

```{r echo=FALSE, message=FALSE, warning=FALSE}
## Analysis
path <- paste0("../data_new/casecrossover/ER")

@@

# read zip pm for counts by zip
smoke_path <- paste0("../data/Oregon_PM/")
zip_pm <- read_csv(paste0(smoke_path, "zip_pm_to_merge_with_acap.csv"))
```

@@

# set list and do cross tabs to find values
# Producing conditional logit model estimates loop 
for(i in 1:length(df_list)){
# dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- colnames(df_to_loop[76])
  # outcome name
  outcome_name <- outcome_list[i]

  # dataframe for analysis creation
  # bind columns back together 
  df_analysis <- df_to_loop %>%
    filter(!is.na(wrf_smk_pm_zip)) %>% 
    # only look at outcomes
    filter(outcome == 1) %>%
    # add another row that makes sure there is a person <15 in the dataframe
    # tricking xtabs to produce a 0 cell for the outcome for age <15
    add_row(outcome = 0, age_ind = 0)

@@

@@

print(print_plot)
# save figure
ggsave("../plot_new/fig_only_er.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")
```

@@

#### Wildfire Smoke and Oregon All Payer All Claims Results
We found a  10 $\mu$g/m^3^ increase in PM~2.5~ was significantly associated with an 2.8% increase for respiratory oucomes  using method of WRF-Chem smoke (OR: 1.028 95% Confidence Interval [CI]: 1.001-1.055), and no significant association for kriging smoke method (OR: 1.018, 95% CI: 0.998-1.038), and also no significant association for GWR smoke method (OR: 1.021, 95% CI: 0.999-1.043)
A 10 $\mu$g/m^3^ increase in PM~2.5~ was significantly associated with an 9.0% increase for asthma outcomes using  WRF-Chem smoke method (OR: 1.090, 95% Confidence Interval [CI]: 1.014-1.171), and 7.9% increase for kriging smoke method (OR: 1.079, 95% CI: 1.036-1.123), and 8.5% increase for GWR smoke method (OR: 1.085, 95% CI: 1.038-1.134)

```{r lags with cdc met adj, warning = F, echo = F, results='asis', eval=FALSE, include = F} 
# note above, results = 'asis' needed to print htmlTables

# looking only at wrf chem, kriging, and geo-weighted; taking out 'Global Smoke' method
method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Global Smoke', 'GWR Smoke')
method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'Geo-Weighted Smoke')

# create an empty list to row bind dataframes together
datalist1 <- list()
datalist2 <- list()

# data wrangling ----
# Producing conditional logit model estimates loop 
for(i in 1:length(df_list)){
 i = 2

  # extract covariates from dataframe (Oregon)
  covariates_df <- df_to_loop[, c(1:26, 71, 74, 76:85)]

  # extract pm values and divide by 10 and ordered
  # which(colnames(df_to_loop)=="wrf_smk_pm_lag1_zip") # code to find column numbers (105)
  # which(colnames(df_to_loop)=="wrf_smk_pm_zip") # code to find column numbers (88)
  # which(colnames(df_to_loop)=="krig_smk_pm_lag1_zip") # code to find column numbers (130)
  # which(colnames(df_to_loop)=="krig_smk_pm_zip") # code to find column numbers (93)
  # which(colnames(df_to_loop)=="geo_smk_pm_lag1_zip") # code to find column numbers (125)
  # which(colnames(df_to_loop)=="geo_smk_pm_zip") # code to find column numbers (92)
  # which(colnames(df_to_loop)=="wrf_temp_lag1_zip") # code to find column numbers (135)
  # which(colnames(df_to_loop)=="wrf_temp_zip") # code to find column numbers (94)

  pm_lag_list <- list(df_to_loop[, c(88, 105:109)]/10, df_to_loop[, c(93, 130:134)]/10,
                  df_to_loop[, c(92, 125:129)]/10, df_to_loop[, c(94, 135:139)]/10)  # create 10 unit increases

# need to think if I need a methods section (probably not since this is it)

  # new loop for each element of the lag list 
  for(k in 1:4){
    k = 1
    # empty matrix (12 x 10 matrix)
    point_estimates <- matrix(nrow = 6, ncol = 10, byrow = T)

    colnames(point_estimates) <- c('outcome', 'pm_method', 'lag', 'n', 'n_events', 
                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')

    # fill in the outcome namedataframe before method loop
    point_estimates[, 1] <- outcome_name
    # repeat the sex category 4 times for each pm method
    point_estimates[, 2] <- method_list[k]

    pm_test <- pm_lag_list[[k]]

    # dataframe for analysis creation
    # bind columns back together 
    df_analysis <- cbind(covariates_df, pm_lag_list[[k]])
#glimpse(df_analysis)

  # loop to run a model for each lag 0-5
    for(j in 0:5){

      j = 0

      # finding locations of columns in dataframe
      # which(colnames(df_analysis)=="daily_meanWS")

      # set column number to evaluate for the pm lag method
      #pm_lag_colnum <- j + 31
      # set column number of cdc temp lag to adjust for
      #temp_lag_colnum <- j + 49
      # set col num for abs temp
      #ab_temp_lag_colnum <- j + 37
      # set col num for relative humidity
      #rh_lag_colnum <- j + 43
      # precip
      #pr_lag_colnum <- j + 55
      # wind speed
      #ws_lag_colnum <- j + 61
      # set row number to fill
      #row_n <- j+1

      # only run the model if the dataframe has observations
      if(nrow(df_analysis) != 0){
      # conditional logistic regression model
     #  mod <- clogit(outcome ~  df_analysis[[pm_lag_colnum]] + df_analysis[[temp_lag_colnum]] + 
     #                df_analysis[[rh_lag_colnum]] + df_analysis[[pr_lag_colnum]] + 
     #                df_analysis[[ws_lag_colnum]] +strata(personkey), df_analysis)

        mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)

      # populate matrix
      point_estimates[row_n, 3] <- j
      point_estimates[row_n, 4] <- mod$n
      point_estimates[row_n, 5] <- mod$nevent
      # odds ratio
      point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)
      # 95% lower bound
      point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
                                        1.96*(summary(mod)$coefficient[1,3])), 3)
      # 95% upper bound
      point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
                                        1.96*(summary(mod)$coefficient[1,3])), 3)
      # standard error
      point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
      # p val
      point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)

      # create else statement that fills matrix with missing so I still have the row
      # in the final dataframe
      } else {point_estimates[row_n, 3] <- method_list[row_n]
              point_estimates[row_n, 4] <- 0
              point_estimates[row_n, 5] <- 0
              point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement

    } # end methods loop

    # combine previous values in dataframe that has all outcome/methods comparisons
    datalist1[[k+1]] <- point_est_df
  } # end sex category loop

  # bind rows of sex category estimates together
 lag_est_df <- bind_rows(datalist1)

 # populate second dataframe list
 datalist2[[i]] <- lag_est_df

} # end of outcome loop

# combine each outcome dataframe itteration in to a big dataset
combined_point_est_df <- bind_rows(datalist2)

#glimpse(combined_point_est_df)
# subset to just geo-smoke estimates for now to reduce number of rows
wrf_geo_lag <- combined_point_est_df %>% 
  # filter columns to just geo-smk
  filter(pm_method == "GWR Smoke" |
         pm_method == "WRF-Chem Smoke") %>% 
  # subset columns I want to put in to the table
  select(2:3, 5:8) %>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
         lower95 = ifelse(lower95 == 99, "--", lower95),
         upper95 = ifelse(upper95 == 99, "--", upper95))

  tab <- htmlTable(txtRound(wrf_geo_lag, digits = 3, 1:3), 
           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes by lagged days",
           # row group by outcome
           rgroup = outcome_list,
           n.rgroup = c(rep(12, 12)), # 6 rows for each lag period for each outcome
           # column headers
           header = c("Est Method", "Lagged Days", "Events",
                      "OR&dagger;", "Lower", "Upper"),
           # column spanner
           cgroup = c("", "95% CI"), 
           n.cgroup = c(4, 2),
           padding.rgroup = "&nbsp;&nbsp;",
           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
           align = "lccccc", # column alignment,
           tfoot="&dagger; Adjusted for lagged CDC temperature, windspeed, relative humidity, and precipitation, accounting for subject."
            ) # end table

# ggplot of odds ratios, facet wrapped by outcomes -----
# convert variables from character to either numeric or factor
# factor preserves the order of the variable
wrf_geo_lag_plot <- combined_point_est_df %>% 
  # filter columns to just geo-smk
  filter(pm_method == "GWR Smoke" |
         pm_method == "WRF-Chem Smoke") %>% 
  # subset columns I want to put in to the table
  select(1:3, 5:8)

# change characters to numeric and factor  
wrf_geo_lag_plot$outcome <- factor(wrf_geo_lag_plot$outcome,
                              levels = unique(wrf_geo_lag_plot$outcome))

wrf_geo_lag_plot$pm_method <- factor(wrf_geo_lag_plot$pm_method,
                                levels = unique(wrf_geo_lag_plot$pm_method))

wrf_geo_lag_plot$lag <- factor(wrf_geo_lag_plot$lag,
                                levels = unique(wrf_geo_lag_plot$lag))

wrf_geo_lag_plot$n_events <- as.numeric(wrf_geo_lag_plot$n_events)
wrf_geo_lag_plot$odds_ratio <- as.numeric(wrf_geo_lag_plot$odds_ratio)
wrf_geo_lag_plot$lower95 <- as.numeric(wrf_geo_lag_plot$lower95)
wrf_geo_lag_plot$upper95 <- as.numeric(wrf_geo_lag_plot$upper95)

## ggplot
  print_plot <- ggplot(wrf_geo_lag_plot,
                  aes(x = lag, y = odds_ratio, color = pm_method), na.rm = T) +
    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
    # custom color 
    scale_color_manual(name = "Smoke-Estimation Method", 
                       values = c("red", "#32115C"),
                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
    facet_wrap(~outcome, nrow = 3, scales = "free") +
    geom_hline(yintercept = 1, linetype=2) +
    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
    #ylim(0.7, 1.4) +
    xlab('Lag Days') +
    # plot theme
    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # strip element
    strip.background = element_rect(colour=NA, fill=NA),
    panel.border = element_rect(fill = NA, color = "black"),
    # facet text size
    strip.text = element_text(size = 10),
    # axis element
    #axis.text.x = element_blank(),
    #axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(angle = 90),
    # legend elements
    legend.position = "bottom")
    #legend.text = element_text(size = 8))

  print(print_plot)

  # save figure
  ggsave("../plot_new/fig4.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

### Time-Stratified by Age Categories

Following table and figure look at some outcomes stratified by age category. Note that kids <=15 likely won't have many outcomes like COPD or MI. I have restricted the dataset to just the geo smoke method to reduce the number of rows/points on the tables and figures to make it easier to read for now. Code can easily be changed to have all PM~2.5~ methods.

For some outcomes, there may be some evidence of effect modificaiton, like COPD, where >65 age group is at significant risk. However, the point estimates are relatively similar for the 15 to 65 strata, just not significant. Some error bars are not shown as they are too wide; usually <15 age category where there are only a handful of outcomes.

```{r time stratified age strata cdc met, warning =F, echo = F, results='asis'} 
# note above, results = 'asis' needed to print htmlTables

# For age categories, I want a loop for each age category. 
# Presenting just WRF and GWR

# looking only at wrf chem, kriging, and geo-weighted; taking out 'Global Smoke' method
method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'GWR Smoke')

# age category list
age_cat_list <- c(0,1,2)

# create an empty list to row bind dataframes together
datalist1 <- list()
datalist2 <- list()

var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')

# dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- var_list[i]
  # outcome name
  outcome_name <- outcome_list[i]

  # extract pm values and divide by 10 and ordered
  #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
  pm_estimates_df <- df_to_loop[, c(88, 93, 92, 94)]/10  # create 10 unit increases

  # new loop for age categories
  for(k in 0:2){

    # empty matrix (12 x 10 matrix)
    point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)

    colnames(point_estimates) <- c('outcome', 'age_cat', 'pm_method', 'n', 'n_events', 
                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')

    # fill in the outcome namedataframe before method loop
    point_estimates[, 1] <- outcome_name
    # repeat the age category 4 times for each pm method
    point_estimates[, 2] <- k

    # dataframe for analysis creation
    # bind columns back together 
    df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
      # remove missing pm values
      filter(!is.na(wrf_smk_pm_zip)) %>% 
      # limit to specific age category
      filter(age_ind == k) %>% 
      # the following code makes sure that the counterfactual values retained are 
      # symetric in that number of obs before = number of obs after
      mutate(obs_diff_admission = (fromdate - date)/7) 
      # dataframe is already for the entire fire season, so I don't need to subset anymore

      # set row number to fill
      row_n <- j-38

      # only run the model if the dataframe has observations
      if(nrow(df_analysis) != 0){
      # conditional logistic regression model
      mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)

      # populate matrix
      point_estimates[row_n, 3] <- method_list[row_n]
      point_estimates[row_n, 4] <- mod$n
      point_estimates[row_n, 5] <- mod$nevent
      # odds ratio
      point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)

      # 95% lower bound
      point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
                                        1.96*(summary(mod)$coefficient[1,3])), 3)
      # 95% upper bound
      point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
                                        1.96*(summary(mod)$coefficient[1,3])), 3)
      # standard error
      point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
      # p val
      point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)

      # create else statement that fills matrix with missing so I still have the row
      # in the final dataframe
      } else {point_estimates[row_n, 3] <- method_list[row_n]
              point_estimates[row_n, 4] <- 0
              point_estimates[row_n, 5] <- 0
              point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement

    } # end methods loop

    # combine previous values in dataframe that has all outcome/methods comparisons
    datalist1[[k+1]] <- point_est_df
  } # end age category loop

  # bind rows of age category estimates together
 age_est_df <- bind_rows(datalist1)

 # populate second dataframe list
 datalist2[[i]] <- age_est_df

} # end of outcome loop

# combine each outcome dataframe itteration in to a big dataset
combined_point_est_df <- bind_rows(datalist2)

wrf_geo_age <- combined_point_est_df %>% 
  # filter columns to just geo-smk
  # filter(pm_method == "GWR Smoke" |
    #      pm_method == "WRF-Chem Smoke") %>% 
  mutate(age_cat2 = ifelse(age_cat == 0, "Less than 15", 
                    ifelse(age_cat == 1, "15-65",
                    ifelse(age_cat == 2, "Greater than 65", NA)))) %>% 
  # subset columns I want to put in to the table
  select(3, 11, 5:8) %>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
         lower95 = ifelse(lower95 == 99, "--", lower95),
         upper95 = ifelse(upper95 == 99, "--", upper95))

  tab <- htmlTable(txtRound(wrf_geo_age, digits = 3, 1:3), 
           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by age",
           # row group by outcome
           rgroup = outcome_list,
           n.rgroup = c(rep(9, 12)), # 6 rows for each age cat for each outcome
           # column headers
           header = c("Est Method", "Age Group", "Events",
                      "OR&dagger;", "Lower", "Upper"),
           # column spanner
           cgroup = c("", "95% CI"), 
           n.cgroup = c(4, 2),
           padding.rgroup = "&nbsp;&nbsp;",
           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
           align = "llcccc", # column alignment,
           tfoot="&dagger; Adjusted for CDC temperature, wind speed, relative humidity, precipitation, and accounting for subject. Time-stratified: referent periods matched to events on same day of week within July to October fire season."
            ) # end table

combined_point_est_df_new <- combined_point_est_df %>%
  mutate(n_events_new = ifelse(as.numeric(n_events)>=100, n_events,
                        ifelse(as.numeric(n_events)<100, "0", NA))) %>%
  select(-n_events) %>%
  rename(n_events = n_events_new)

# ggplot of odds ratios, facet wrapped by outcomes -----
# convert variables from character to either numeric or factor
# factor preserves the order of the variable
wrf_geo_age_plot <- combined_point_est_df_new %>% 
  # filter columns to just geo-smk
 #  filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
  # do not plot results with less than 15 events, create missing vals
  # in mutate
  mutate(odds_ratio = ifelse(as.numeric(n_events) < 15, 
                             99, odds_ratio),
         lower95 = ifelse(as.numeric(n_events) < 15, 
                             99, lower95),
         upper95 = ifelse(as.numeric(n_events) < 15, 
                             99, upper95),
         age_cat2 = ifelse(age_cat == 0, "<15", 
                    ifelse(age_cat == 1, "15-65",
                    ifelse(age_cat == 2, ">65", NA)))) %>% 
  # subset columns I want to put in to the table
  select(1, 3, 11, 10, 5:7) %>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
         lower95 = ifelse(lower95 == 99, NA, lower95),
         upper95 = ifelse(upper95 == 99, NA, upper95))

# change characters to numeric and factor  
wrf_geo_age_plot$outcome <- factor(wrf_geo_age_plot$outcome,
                              levels = unique(wrf_geo_age_plot$outcome))

wrf_geo_age_plot$pm_method <- factor(wrf_geo_age_plot$pm_method,
                                levels = unique(wrf_geo_age_plot$pm_method))

wrf_geo_age_plot$age_cat2 <- factor(wrf_geo_age_plot$age_cat2,
                                levels = unique(wrf_geo_age_plot$age_cat2))

wrf_geo_age_plot$n_events <- as.numeric(wrf_geo_age_plot$n_events)
wrf_geo_age_plot$odds_ratio <- as.numeric(wrf_geo_age_plot$odds_ratio)
wrf_geo_age_plot$lower95 <- as.numeric(wrf_geo_age_plot$lower95)
wrf_geo_age_plot$upper95 <- as.numeric(wrf_geo_age_plot$upper95)

## ggplot 
  print_plot <- ggplot(wrf_geo_age_plot,
    aes(x = age_cat2, y = odds_ratio, color = pm_method), na.rm = T) +
    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
    # custom color 
    scale_color_manual(name = "Smoke-Estimation Method", 
                       values = c("red",  "blue", "#32115C"),
                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
    facet_wrap(~outcome, nrow = 3, scales = "free") +
    geom_hline(yintercept = 1, linetype=2) +
    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
    #ylim(0, 2) +
    xlab('Age Category ') +
    # plot theme
    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # strip element
    strip.background = element_rect(colour=NA, fill=NA),
    panel.border = element_rect(fill = NA, color = "black"),
    # facet text size
    strip.text = element_text(size = 10),
    # axis element
    #axis.text.x = element_blank(),
    #axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(angle = 90),
    # legend elements
    legend.position = "bottom")
    #legend.text = element_text(size = 8))

  print(print_plot)
  # save figure
  ggsave("../plot_new/age_fig_three.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")

### Seasonal Time-Stratified by Sex

Association with smoke PM~2.5~ and health outcomes stratified by sex. May be some evidence of effect modification by sex, particularly where males appear to be at greater risk for certain CVD outcomes with increasing smoke. However, only marginally significant. Also, broken arm estimate seems a little strange, suggesting females at greater risk too. Larger sample size may help to answer this.

```{r sex strata cdc met adj, warning = F, echo = F, results='asis'} 
# note above, results = 'asis' needed to print htmlTables

# For age categories, I want a loop for each age category.

var_list <- c('respiratory', 'asthma', 'pneumonia',  'acute_bronch', 'copd', 
              'cvd', 'isch_heart_dis', 'arrhythmia', 'heart_failure', 
              'cerbrovas_dis', 'myocardial_infarc', 'broken_arm')

# looking only at wrf chem, kriging, and geo-weighted; taking out 'Global Smoke' method
method_list <- c('WRF-Chem Smoke', 'Kriging Smoke', 'GWR Smoke')

# sex category list
sex_strata_list <- c(0,1)

# create an empty list to row bind dataframes together
datalist1 <- list()
datalist2 <- list()

  # dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- var_list[i]
  # outcome name
  outcome_name <- outcome_list[i]

  # extract pm values and divide by 10 and ordered
  #which(colnames(df_to_loop)=="global_smk_pm_zip") # code to find column numbers
  pm_estimates_df <- df_to_loop[, c(88, 93, 92, 94)]/10  # create 10 unit increases

  # new loop for sex categories
  for(k in 0:1){

    # empty matrix (12 x 10 matrix)
    point_estimates <- matrix(nrow = 3, ncol = 10, byrow = T)

    colnames(point_estimates) <- c('outcome', 'sex', 'pm_method', 'n', 'n_events', 
                                   'odds_ratio', 'lower95', 'upper95', 'se', 'p_val')

    # fill in the outcome namedataframe before method loop
    point_estimates[, 1] <- outcome_name
    # repeat the sex category 4 times for each pm method
    point_estimates[, 2] <- k

    # dataframe for analysis creation
    # bind columns back together 
    df_analysis <- cbind(covariates_df, pm_estimates_df) %>% 
      # remove missing pm values
      filter(!is.na(wrf_smk_pm_zip)) %>% 
      # limit to specific sex category
      filter(sex_ind == k) %>% 
      # the following code makes sure that the counterfactual values retained are 
      # symetric in that number of obs before = number of obs after
      mutate(obs_diff_admission = (fromdate - date)/7) 
      # dataframe is already for the entire fire season, so I don't need to subset anymore

      # set row number to fill
      row_n <- j-38

      # only run the model if the dataframe has observations
      if(nrow(df_analysis) != 0){
      # conditional logistic regression model
      mod <- clogit(outcome ~ df_analysis[[j]] + wrf_temp_zip + strata(personkey), df_analysis)

      # populate matrix
      point_estimates[row_n, 3] <- method_list[row_n]
      point_estimates[row_n, 4] <- mod$n
      point_estimates[row_n, 5] <- mod$nevent
      # odds ratio
      point_estimates[row_n, 6] <- round(exp(summary(mod)$coefficient[1,1]), 3)

      # 95% lower bound
      point_estimates[row_n, 7] <- round(exp((summary(mod)$coefficient[1,1]) -
                                        1.96*(summary(mod)$coefficient[1,3])), 3)
      # 95% upper bound
      point_estimates[row_n, 8] <- round(exp((summary(mod)$coefficient[1,1]) +
                                        1.96*(summary(mod)$coefficient[1,3])), 3)
      # standard error
      point_estimates[row_n, 9] <- round(summary(mod)$coefficient[1,3], 4)
      # p val
      point_estimates[row_n, 10] <- round(summary(mod)$coefficient[1,5], 4)

      # create else statement that fills matrix with missing so I still have the row
      # in the final dataframe
      } else {point_estimates[row_n, 3] <- method_list[row_n]
              point_estimates[row_n, 4] <- 0
              point_estimates[row_n, 5] <- 0
              point_estimates[row_n, c(6:10)] <- 99 } # end 'if else' statement

    } # end methods loop

    # combine previous values in dataframe that has all outcome/methods comparisons
    datalist1[[k+1]] <- point_est_df
  } # end sex category loop

  # bind rows of sex category estimates together
 sex_est_df <- bind_rows(datalist1)

 # populate second dataframe list
 datalist2[[i]] <- sex_est_df

} # end of outcome loop

# combine each outcome dataframe itteration in to a big dataset
combined_point_est_df <- bind_rows(datalist2)

wrf_geo_sex <- combined_point_est_df %>% 
  # filter columns to just geo-smk
  # filter(pm_method == "GWR Smoke" | pm_method == "WRF-Chem Smoke") %>% 
  mutate(sex_cat = ifelse(sex == 0, "Male", 
               ifelse(sex == 1, "Female", NA))) %>% 
  # subset columns I want to put in to the table
  select(3, 11, 5:8)%>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, "--", odds_ratio),
         lower95 = ifelse(lower95 == 99, "--", lower95),
         upper95 = ifelse(upper95 == 99, "--", upper95))

  tab <- htmlTable(txtRound(wrf_geo_sex, digits = 3, 1:3), 
           caption = "Association between a 10 ug/m^3 in PM2.5 geo smoke and health outcomes stratified by sex",
           # row group by outcome
           rgroup = outcome_list,
           n.rgroup = c(rep(6, 12)), # 2 rows for each sex strata for each outcome
           # column headers
           header = c("Est Method", "Sex Strata", "Events",
                      "OR&dagger;", "Lower", "Upper"),
           # column spanner
           cgroup = c("", "95% CI"), 
           n.cgroup = c(4, 2),
           padding.rgroup = "&nbsp;&nbsp;",
           css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
           align = "llcccc", # column alignment,
           tfoot="&dagger; Adjusted for temperature, accounting for subject. Time-stratified: referent periods matched to events on same day of week within July to October fire season."
            ) # end table

# ggplot of odds ratios, facet wrapped by outcomes -----
# convert variables from character to either numeric or factor
# factor preserves the order of the variable
wrf_geo_sex_plot <- combined_point_est_df %>% 
  # filter columns to just geo-smk
 #  filter(pm_method == "GWR Smoke" | 
   #       pm_method == "WRF-Chem Smoke") %>% 
  mutate(sex_cat = ifelse(sex == 0, "Male", 
               ifelse(sex == 1, "Female", NA))) %>%
  # subset columns I want to put in to the table
  select(1, 3, 11, 5:8) %>% 
  mutate(odds_ratio = ifelse(odds_ratio == 99, NA, odds_ratio),
         lower95 = ifelse(lower95 == 99, NA, lower95),
         upper95 = ifelse(upper95 == 99, NA, upper95))

# change characters to numeric and factor  
wrf_geo_sex_plot$outcome <- factor(wrf_geo_sex_plot$outcome,
                              levels = unique(wrf_geo_sex_plot$outcome))

wrf_geo_sex_plot$pm_method <- factor(wrf_geo_sex_plot$pm_method,
                                levels = unique(wrf_geo_sex_plot$pm_method))

wrf_geo_sex_plot$sex_cat <- factor(wrf_geo_sex_plot$sex_cat,
                                levels = unique(wrf_geo_sex_plot$sex_cat))

wrf_geo_sex_plot$n_events <- as.numeric(wrf_geo_sex_plot$n_events)
wrf_geo_sex_plot$odds_ratio <- as.numeric(wrf_geo_sex_plot$odds_ratio)
wrf_geo_sex_plot$lower95 <- as.numeric(wrf_geo_sex_plot$lower95)
wrf_geo_sex_plot$upper95 <- as.numeric(wrf_geo_sex_plot$upper95)

## ggplot
  print_plot <- ggplot(wrf_geo_sex_plot,
    aes(x = sex_cat, y = odds_ratio, color = pm_method), na.rm = T) +
    geom_point(position = position_dodge(width = 0.5), na.rm = T) + 
    geom_errorbar(aes(ymin=lower95, ymax=upper95), 
                  position = position_dodge(width = 0.5), width = 0.2, na.rm =T) +
    # custom color 
    scale_color_manual(name = "Smoke-Estimation Method", 
                       values = c("red", "blue", "#32115C"),
                       guide = guide_legend(title.position = "top", title.hjust = 0.5)) +
    facet_wrap(~outcome, nrow = 3, scales = "free") +
    geom_hline(yintercept = 1, linetype=2) +
    #ggtitle('Association Between PM2.5 from \n Wildfire Smoke on Hospitalizations') +
    ylab(expression(paste("Odds Ratio for 10µg/m"^3, " Increase in PM"[2.5]))) +
    #ylim(0, 2) +
    xlab('Sex') +
    # plot theme
    theme(panel.background = element_rect(fill = 'white', colour = 'black'),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # strip element
    strip.background = element_rect(colour=NA, fill=NA),
    panel.border = element_rect(fill = NA, color = "black"),
    # facet text size
    strip.text = element_text(size = 10),
    # axis element
    #axis.text.x = element_blank(),
    #axis.title.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.y = element_text(angle = 90),
    # legend elements
    legend.position = "bottom")
    #legend.text = element_text(size = 8))

  print(print_plot)
  # save figure
  ggsave("../plot_new/sex_fig_three.pdf", plot = print_plot, 
       width = 12, height = 8, units = "in")
